---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# doctest


<!-- badges: start -->
<!-- badges: end -->

Doctests are documentation combined with tests  The doctest package 
helps you write doctests by adding tags to your [roxygen](https://roxygen2.r-lib.org/) documentation.


## Installation

You can install the development version of doctest like so:

``` r
devtools::install("hughjonesd/doctest")
```

## Example

Here's some documentation for a function:

```r
```{r child = "safe-mean.Rmd"}
```
```

This will create tests like:

```{r, echo = FALSE, warning = FALSE}
library(doctest)
library(roxygen2)
text <- readLines("safe-mean.Rmd")
results <- roc_proc_text(doctest(), text)
roclet_output(doctest(), results)
```

To use testygen, change your package DESCRIPTION to add the `doctest`
roclet to roxygen, like this:

```
Roxygen: list(roclets = c("collate", "rd", "namespace", "doctest")) 
```
