---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# doctest


<!-- badges: start -->
<!-- badges: end -->

Doctests are documentation combined with tests  The doctest package 
helps you write doctests by adding tags to your [roxygen](https://roxygen2.r-lib.org/) documentation.


## Example

Here's some documentation for a function:

```r
```{r child = "safe-mean.Rmd"}
```
```

This will create tests like:

```{r, echo = FALSE, warning = FALSE}
library(doctest)
library(roxygen2)
text <- readLines("safe-mean.Rmd")
results <- roc_proc_text(doctest(), text)
roclet_output(doctest(), results)
```


## Usage

To use doctest, put a line in your package DESCRIPTION to add the `doctest`
roclet to roxygen, like this:

```
Roxygen: list(roclets = c("collate", "rd", "namespace", "doctest::doctest")) 
```

Then in the package directory, run:

```r
library(doctest)
roxygen2::roxygenize()
```

At present, you can't use doctest from the RStudio keyboard shortcut
`Ctrl + Shift + D`, because this always uses the standard roxygen2 roclets. 
However, you can bind the RStudio addin "Devtools: document a package" to a
keyboard shortcut. This will use the roclets from your package DESCRIPTION file.

## doctest tags

The doctest package adds four tags to roxygen:

* `@expect` takes the next expression and writes an expectation for it.

```r
#'
#' @expect equals(2)
1+1
```

You can use any `expect_*` function from `testthat`. Omit the `expect_` at the
start.

* `@test *test-name*` creates a new test.

```r
#' @test Should fail on NA
#' @expect error()
#' my_func(NA)
```

By default, all expectations are created in a single test.

* `@skiptest` omits lines of the example from the test, and

* `@unskip` stops omitting lines

```r
#' @skiptest
#' # No need to test plotting
#' plot(1:10, my_func(1:10))
#' @unskip
```

## Installation

You can install the development version of doctest like so:

``` r
devtools::install("hughjonesd/doctest")
```
