---
title: "doctest"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
vignette: >
  %\VignetteIndexEntry{doctest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction to doctest

```{r, child="../man/rmd/intro.Rmd"}
```

```{r, child="../man/rmd/example.Rmd"}
```

```{r, child="../man/rmd/usage.Rmd"}
```

Here's a simple workflow to start using doctest:

1. Alter your package DESCRIPTION as above.

2. In your roxygen documentation, replace `@examples` by `@doctest`.

3. In the package directory run `roxygen2::roxygenize()` 
or `devtools::document()` to create documentation. You should see Rd files created as normal in the `man/` directory, including `\examples` sections.

4. Add `@expect` tags to your `@doctest` sections.

5. Run `roxygenize()` again. You will now see new files created in 
the `tests/testthat` directory, with the name `test-doctest-<topic name>.R`.

6. Run `devtools::test()` and check that your tests pass. 

At present, you can't use doctest from the RStudio keyboard shortcut
`Ctrl + Shift + D`, because this always uses the standard roxygen2 roclets. 
However, you can bind the RStudio addin "Devtools: document a package" to a
keyboard shortcut. This will use the roclets from your package DESCRIPTION 
file.


## Tags

The doctest package adds these tags to roxygen:


### `@doctest`

Use `@doctest` instead of `@examples`: 

```r
#' @doctest
#' 
#' # ... examples for your function
```

The content of `@doctest` will be used in the .Rd "examples" section, and in a testthat test.

You can have more than one `@doctest` section. Each section creates one test
like `test_that("Test name", {...})`. You can name the doctest, or leave it blank
for a default name. All the sections will be merged into a single .Rd example.


```r
#' @doctest Positive numbers
#' x <- 1
#' @expect equal(x)
#' abs(x)
#'
#' @doctest Negative numbers
#' x <- -1
#' @expect equal(-x)
#' abs(x)
```


### `@expect`

`@expect` writes a testthat expectation.

```r
#' @expect equal(4)
#' 2 + 2

```

You can use any `expect_*` function from `testthat`. Omit the `expect_` at the
start of the call.

The expression on the next line will be substituted as the first argument
into the `expect` call:

```r
expect_equal(2 + 2, 4)
```

Use a dot `.` to substitute in different places:

```r
#' @expect equal(., rev(.))
#' c("T", "E", "N", "E", "T")
```

This becomes:

```r
expect_equal(c("T", "E", "N", "E", "T"), rev(c("T", "E", "N", "E", "T")))
```

### `@expectRaw`

`@expectRaw` writes an expectation, without substituting the next expression:

```r
#' x <- 2 + 2
#' @expectRaw equal(x, 4)
```


### `@snap`

`@snap` is shorthand for `@expect snapshot()`. This creates a 
[snapshot test](https://testthat.r-lib.org/articles/snapshotting.html),
which is useful for checking that complex examples haven't changed.


### `@omit` and `@resume`

By default, the doctest uses the whole example. `@omit` stops including the
example in the doctest until the next expectation or other tag. You can use
`@resume` to restart including lines without creating a new expectation.

```r
#' myfunc(1)
#' 
#' @omit
#' # No need to test plotting
#' plot(1:10, my_func(1:10))
#' 
#' @resume
#' x <- NA
#' @expect warning()
#' myfunc(x)
```

## Writing good doctests

Tests and documentation are similar, but not identical. Tests need to cover
difficult corner cases. Examples need to convey the basics to the user. 
I like the following advice:

> ... write the best possible documentation, and [R] makes sure the code
> samples in your documentation actually compile and run [and do what they
> are supposed to do]

*Programming Rust*, Blandy, Orendorff and Tindall, 2021

In particular, use doctest as an *addition* to manually created tests, not a 
*substitute* for them. Use doctest for relatively simple tests of basic functionality. If it's hard to specify what to test for, consider using
`@snap` to capture output. For more complex test cases, write a test file manually.

```r
#' @snap
summary(model)
```



## Related packages

The [roxytest](https://mikldk.github.io/roxytest/) and
[roxut](https://github.com/bryanhanson/roxut) packages both allow you to write
tests in roxygen blocks. Doctest is slightly different because it combines tests
with examples. The [exampletestr](https://github.com/rorynolan/exampletestr/)
package uses roxygen examples to generate a test skeleton which you can fill in
yourself.



